Locally Conservative Discontinuous Petrov-Galerkin Finite Elements for Fluid Problems

### 2013-02-25 docs/Truman/LocalConservation/paper.tex
Wrote up Erickson and skewed convection-diffusion.

\documentclass[letterpaper]{article}
\usepackage[margin=1.0in]{geometry}
\usepackage{authblk}
\usepackage{amsmath, amssymb, amsthm, mathtools}
\usepackage{graphicx}
\usepackage{subfigure}

@@

\section{Numerical Experiments}

% A cursory look at \eqnref{confusionBF} tells us which function spaces to look
% in for each variable:
% \[
% \begin{array}{lll}
% u\in L^2(\Omega_h) \quad& \hat u\in H^{1/2}(\Gamma_h) \quad& v\in H^1(\Omega_h)\\
% \bsigma\in \mathbf{L}^2(\Omega_h) \quad& \hat t\in H^{-1/2}(\Gamma_h) \quad&
% \btau\in \bs{H}(\bs{div},\Omega_h)\,.\\
% \end{array}
% \]

% All that is left to pin down this problem is a definition of our test norm so
% we can invert the Riesz operator and calculate our optimal test functions.
% Within each element, we perform a Bubnov-Galerkin solve for the optimal test
% functions. Define finite-dimensional subspaces
% $\mb{U}_h\subset\mb{U}:=L^2(\Omega_h)\times\mb{L}^2(\Omega_h)\times
% H^{1/2}(\Gamma_h)\times H^{-1/2}(\Gamma_h)$ the trial space and 
% $\mb{V}_h\subset\mb{V}:=H^1(\Omega_h)\times H(div,\Omega_h)$ the ``enriched''
% test space.
% For each $\mathbf{u}=\{u,\bsigma,\hat u,\hat t\}\in\mathbf{U}_h$, find
% $\mathbf{v}_{\mathbf{u}}=\{v_\mathbf{u},\btau_\mathbf{u}\}\in\mathbf{V_h}$ such that
% \[
% (\mathbf{v_u},\mathbf{w})_\mathbf{V}=b(\mathbf{u},\mathbf{w})\quad\forall\mathbf{w}\in\mathbf{V}
% \]
% As mentioned earlier the choice of test norm on $V$ can have profound
% influence on the robustness of our method. Unfortunately, the structure of the
% optimal test norm makes it non-localizable. For many problems, it suffices to
% use the so-call quasi-optimal test norm which is based on the adjoint of the
% $B$ operator, but for convection-diffusion type equations, the adjoint
% develops boundary layers which make solving for the optimal test functions
% much more difficult for small diffusion. In an earlier consideration of DPG
% for convection-diffusion problems, Chan \etal developed the more robust
% test norm\cite{ChanHeuerThanhDemkowicz2012}, 
% \begin{equation}
% \label{eq:robustNorm}
% \norm{(v,\btau)}^2_{\mathbf{V},\Omega_h}=
% \norm{\nabla\cdot\btau}^2+\norm{\min\left\{\frac{1}{\sqrt{\epsilon}},\frac{1}{\sqrt{|K|}}\right\}\btau}^2
% +\epsilon\norm{\nabla v}^2+\norm{\bbeta\cdot\nabla v}^2
% +\norm{\min\left\{\sqrt{\frac{\epsilon}{|K|}},1\right\}v^2}\,.
% \end{equation}
% Unfortunately, this test norm also has a few issues. For one, some of the
% assumptions that went into its development break down as the flow field
% degenerates to zero. In such cases the final $L^2$ term on $v$ can register
% higher error levels and trigger unnecessary refinements in smooth regions, see
% ref{}.
%TODO: have Jesse write about this

% \subsection{Locally Conservative Formulation}
% A simple control volume analysis will tell us that a locally conservative
% method must enforce that
% \begin{equation}
% \int_{\partial K}\hat t=\int_K f\,,\quad\forall K\in\Omega_h\,,
% \label{eq:localconservation}
% \end{equation}
% which is equivalent to having the set
% $\mathbf{v}_K:=\{v,\btau\}=\{1_K,\boldsymbol0\}$ for $K=1,\dots,N$ ($N$ is the
% number of mesh elements)
% in the test space, where each $1_K$ has value one on element $K$ and zero
% elsewhere.  In fact, if we insert this test function into
% \eqnref{confusionBF}, all of the $\tau$ and $\grad v$ terms vanish and we are
% left exactly with this condition.
% Numerical experiments imply that local conservation occurs in one dimension,
% but the standard DPG method is not exactly locally conservative for higher
% dimensional problems.
% 
% Following Moro \etal\cite{MoroNguyenPeraire11}, we can explicitly augment our
% test space with constants through the use of Lagrange multipliers. Going back
% to \eqnref{rieszapplied}, we can define our Lagrange function,
% \begin{equation}
% \label{eq:lagrangian}
% L(u_h,\blambda) = \frac{1}{2}\norm{R_V^{-1}(Bu_h-l)}_V^2
% -\sum_K\lambda_K\underbrace{\langle Bu_h-l,\mathbf{v}_K\rangle}_
% {\langle\hat t, 1_K\rangle_{\partial K}-\langle g,1_K\rangle_K}\,,
% \end{equation}
% where $\blambda=\{\lambda_1,\cdots,\lambda_N\}$.
% We then proceed as before and find the critical points of \eqnref{lagrangian},
% \begin{equation}
% \label{eq:modifiedBF}
% \frac{\partial L(u_h,\blambda)}{\partial u_h}=b(u_h,R_V^{-1}B\delta u_h)
% -l(R_V^{-1}B\delta u_h)
% -\sum_K\lambda_K b(\delta
% u_h,\mathbf{v}_K)=0\,,\quad\forall\delta u_h\in U_h
% \end{equation}
% \begin{equation}
% \label{eq:constraint}
% \frac{\partial
% L(u_h,\blambda)}{\partial\lambda_K}=-b(u_h,\mathbf{v}_K)+l(\mathbf{v}_K)=0\,,\quad\forall
% K\,.
% \end{equation}
% Equation \eqnref{modifiedBF} is just \eqnref{DPGmethod} with the extra
% Lagrange terms. As usual, the second equation just enforces the constraint. As
% a consequence, we now explicitly have constants in our test space and should
% enforce local conservation to machine precision. The negative side is that we
% have added an additional unknown associated with every mesh element and turned
% our well-behaved minimization into a saddle point problem.
% 
% This change has further consequences to how we compute our optimal test
% functions. Since constants are now explicitly represented in the test space,
% we only need to search for optimal test functions in the orthogonal complement
% of constants. The final term in \eqnref{robustNorm} is somewhat troublesome,
% but becomes unnecessary when solving in the orthogonal complement of
% constants. Instead, we can replace it with a much nicer zero mean term. Thus,
% \eqnref{robustNorm} becomes
% \begin{equation}
% \norm{(v,\btau)}^2_{\mathbf{V},\Omega_h}=
% \norm{\nabla\cdot\btau}^2+\norm{\min\left\{\frac{1}{\sqrt{\epsilon}},\frac{1}{\sqrt{|K|}}\right\}\btau}^2\\
% +\epsilon\norm{\nabla v}^2+\norm{\bbeta\cdot\nabla
% v}^2+\left(\int_Kv\right)^2\,.
% \end{equation}

>>>>

\documentclass[letterpaper]{article}
\usepackage[margin=1.0in]{geometry}
\usepackage{authblk}
\usepackage{amsmath, amssymb, amsthm, mathtools}
\usepackage{graphicx}
% \usepackage{subfigure}

@@

\section{Numerical Experiments}
We explore each of the following experiments with two sets of test norms: the
standard graph norm, and the aforementioned robust norm.

\subsection{Erickson-Johnson Model Problem}
The Erickson-Johnson problem is one of the few convection-diffusion problems
with a known analytical solution. Take a domain $\Omega=[0,1]^2$, with
$\bbeta=(1,0)^T$, $f=0$ and boundary conditions $\hat t=u_0$ when
$\beta_n\le0$, and $u_0$ is the trace of the exact solution. For
$n=1,2,\cdots$, let
$\lambda_n=n^2*\pi^2*\epsilon$,
$r_n=\frac{1+\sqrt{1+4\epsilon\lambda_n}}{2\epsilon}$, and 
$s_n=\frac{1-\sqrt{1+4\epsilon\lambda_n}}{2\epsilon}$, the exact solution
is
\begin{equation}
u(x,y)=C_0+\sum_{n=1}^\infty C_n\frac{\exp(s_n(x-1))-\exp(r_n(x-1))}
{r_n\exp(-s_n)-s_n\exp(-s_n)}\cos(n\pi y)\,.
\label{eq:ericksonExact}
\end{equation}
Taking $\epsilon=0.01$, $C_1=1$, and $C_{n\neq1}=0$, the exact solution looks
like Figure \ref{fig:erickson}. We solve this problem with both conservative
and nonconservative formulations of DPG with both graph and robust test norms.

\begin{figure}[h!]
\centering
\includegraphics[width=0.6\textwidth]{figs/Erickson/exact.png}
\caption{Erickson-Johnson exact solution}
\label{fig:erickson}
\end{figure}

We plot the $L^2$ and energy error for the various methods in Figures
\ref{fig:ericksonGraphError} and \ref{fig:ericksonRobustError}. Encouragingly,
the conservative and nonconservative plots appear to lay right on top of each
other. This indicates that the conservative formulation at least doensn't hurt
our convergence significantly.

\begin{figure}[h!]
\centering
\parbox{0.45\textwidth}{
\includegraphics[width=0.45\textwidth]{figs/Erickson/graphError.pdf}
\caption{Error in Erickson-Johnson solution with graph norm}
\label{fig:ericksonGraphError}
}\quad
\parbox{0.45\textwidth}{
\includegraphics[width=0.45\textwidth]{figs/Erickson/robustError.pdf}
\caption{Error in Erickson-Johnson solution with robust norm}
\label{fig:ericksonRobustError}
}
\end{figure}

We plot the flux imbalance errors in Figures \ref{fig:ericksonGraphFlux} and
\ref{fig:ericksonRobustFlux}. As expected, the element conservative
formulation brings these errors to the order of machine precision, but it is
worth pointing out that the flux imbalances for standard DPG are reasonable,
especially under refinement.

\begin{figure}[h!]
\centering
\parbox{0.45\textwidth}{
\includegraphics[width=0.45\textwidth]{figs/Erickson/graphFlux.pdf}
\caption{Flux imbalance in Erickson-Johnson solutions with graph norm}
\label{fig:ericksonGraphFlux}
}\quad
\parbox{0.45\textwidth}{
\includegraphics[width=0.45\textwidth]{figs/Erickson/robustFlux.pdf}
\caption{Flux imbalance in Erickson-Johnson solutions with robust norm}
\label{fig:ericksonRobustFlux}
}
\end{figure}

\subsection{Skewed Convection-Diffusion Problem}
This is a standard convection-diffusion problem with a skewed convection
vector relative to the mesh. The problem setup is identical to the
Erickson-Johnson problems except that $\bbeta=(2,1)$. The following results
are for $\epsilon=10^{-4}$.

On our initial mesh, we see no noticeable difference between conservative and
nonconservative results, though the choice of test norm affects things
dramatically. The graph norm produces a solution which is comparable to an L2
project of the exact solution. In this case, the field variable don't show the
very small boundary layer along the right and top surfaces, though a
visualization of the traces would show that the wall boundary condition is in
fact being enforced. As we refine, the field variables will begin to reflect
the boundary layer. On the other hand, the robust norm does show a boundary
layer being formed in the field variables.

\begin{figure}[h!]
\centering
\parbox{0.45\textwidth}{
\includegraphics[width=0.45\textwidth]{figs/Confusion/graph0c.png}
\caption{Conservative skewed convection-diffusion with graph norm, initial mesh}
\label{fig:confusionGraph0c}
}\quad
\parbox{0.45\textwidth}{
\includegraphics[width=0.45\textwidth]{figs/Confusion/robust0c.png}
\caption{Conservative skewed convection-diffusion with robust norm, initial mesh}
\label{fig:confusionRobust0c}
}
\end{figure}

As we refine, the three results, nonconservative with graph or robust norm,
and conservative with graph norm converge to a an essentially identical
solution (to the eyeball norm), as represented by
Figure~\ref{fig:confusionGraph8c}. However, the conservative formulation with
the robust norm seems to exhibit some noise at the right wall as seen in
Figure~\ref{fig:confusionRobust8c}.

% \begin{figure}[h!]
% \centering
% \parbox{0.45\textwidth}{
% \includegraphics[width=0.45\textwidth]{figs/Confusion/graph8nc.png}
% \caption{Nonconservative skewed convection-diffusion with graph norm, after 8
% refinements}
% \label{fig:confusionGraph8nc}
% }\quad
% \parbox{0.45\textwidth}{
% \includegraphics[width=0.45\textwidth]{figs/Confusion/robust8nc.png}
% \caption{Nonconservative skewed convection-diffusion with robust norm, after 8
% refinements}
% \label{fig:confusionRobust8nc}
% }
% \end{figure}

\begin{figure}[h!]
\centering
\parbox{0.45\textwidth}{
\includegraphics[width=0.45\textwidth]{figs/Confusion/graph8c.png}
\caption{Conservative skewed convection-diffusion with graph norm, after 8
refinements}
\label{fig:confusionGraph8c}
}\quad
\parbox{0.45\textwidth}{
\includegraphics[width=0.45\textwidth]{figs/Confusion/robust8c.png}
\caption{Conservative skewed convection-diffusion with robust norm, after 8
refinements}
\label{fig:confusionRobust8c}
}
\end{figure}

If we zoom in on one of the noisy areas, as we do in
Figure~\ref{fig:confusionRobust8c_zoom}, we see overshoots on the order of 32
and undershoots on the order of -13, where the exact solution varies between 0
and 1. The cause of these overshoots and undershoots is a little mystery,
especially because they don't show up until 8 refinement steps in. We will
explore this problem a little more when we look at the wedge problem.

\begin{figure}[h!]
\centering
\includegraphics[width=0.6\textwidth]{figs/Confusion/robust8c_zoom.png}
\caption{Zoomed in conservative skewed convection-diffusion problem after 8
refinements}
\label{fig:confusionRobust8c_zoom}
\end{figure}

Finally, we plot the flux imbalances for the several methods in
Figure~\ref{fig:confusionGraphFlux} and Figure~\ref{fig:confusionRobustFlux}.
Interestingly, it appears the effectiveness of our local conservation
formulation drops as we refine with the graph norm.

\begin{figure}[h!]
\centering
\parbox{0.45\textwidth}{
\includegraphics[width=0.45\textwidth]{figs/Confusion/graphFlux.pdf}
\caption{Flux imbalance in skewed confusion solutions with graph norm}
\label{fig:confusionGraphFlux}
}\quad
\parbox{0.45\textwidth}{
\includegraphics[width=0.45\textwidth]{figs/Confusion/robustFlux.pdf}
\caption{Flux imbalance in skewed confusion solutions with robust norm}
\label{fig:confusionRobustFlux}
}
\end{figure}
